Ensemble Stability of Biological Networks

### 2014-09-23 paper.tex
added reference for graph edit distance

\todo{add reference for graph edit distance}
The relationship between $G$ and $\hier(G)$ for all $G$ with a given number of vertices suggests a heuristic method of quantifying the degree of hierarchy of a given graph and thus of the system structure it represents. The most hierarchical system is considered to be the graph corresponding to the total ordering, which for three nodes is given in \ref{fig:scc}A. This graph maximizes the number of links between strongly connected components, which also implies maximizing the number of strongly connected components. The graph edit distance on a fixed number of vertices from one graph to another is defined as the minimum number of modifications of the first graph in order to transform it into the second. This distance between any given graph and the total ordering thus quantitatively represents how far a graph is from being perfectly hierarchical. In this work we take this distance to be the definition of hierarchy.

@@

Since every acyclic digraph can be embedded into a totally ordered set, we may assume without loss of generality that our components have been ordered in a way such that, if $(i,j) \in \hier(G)$, then $i < j$.  Hence, after substituting into our expression and simplifying using the identity
$$\sum_{i=1}^{n-1}\sum_{j=i+1}^{n}v_i v_j~=~\frac{1}{2} \left( \sum_{i=1}^{n} \right)^2-\frac{1}{2} \sum_{i=1}^{n} v_i^2,$$
we arrive at the following upper bound on robustness in terms of the connected components:
\begin{align}
R_{\mathrm{max}}(C_1, \ldots C_n) =
\frac{\frac{1}{2} ((\sum_i n_i)^2 - \sum_i n_i^2) + \sum_i \connectivity_i R(C_i)}
{\frac{1}{2} ((\sum_i v_i)^2 - \sum_i v_i^2) + \sum_i \connectivity_i}, \label{eq:sccmaxrobustness}
\end{align}
Furthermore, this bound is attained.  Suppose that $G_{\mathrm{tot}}$ is the graph on $n$ nodes with a link from node $i$ to node $j$ whenever $i < j$.  Then, by the construction described in the previous section on network hierachy, we have a graph $G_{\mathrm{max}}$ such that the components of $G_{\mathrm{max}}$ are $C_1, \ldots C_n$ and $\hier (G_{\mathrm{max}}) = G_{\mathrm{tot}}$.  By our formula, $R(G_{\mathrm{max}}) = R_{\mathrm{max}}(C_1, \ldots C_n)$.

>>>>

The relationship between $G$ and $\hier(G)$ for all $G$ with a given number of vertices suggests a heuristic method of quantifying the degree of hierarchy of a given graph and thus of the system structure it represents. The most hierarchical system is considered to be the graph corresponding to the total ordering, which for three nodes is given in \ref{fig:scc}A. This graph maximizes the number of links between strongly connected components, which also implies maximizing the number of strongly connected components. The graph edit distance on a fixed number of vertices from one graph to another is defined as the minimum number of modifications of the first graph in order to transform it into the second \cite{Axenovich2011}. This distance between any given graph and the total ordering thus quantitatively represents how far a graph is from being perfectly hierarchical. In this work we take this distance to be the definition of hierarchy.

@@

Since every acyclic digraph can be embedded into a totally ordered set, we may assume without loss of generality that our components have been ordered in a way such that, if $(i,j) \in \hier(G)$, then $i < j$.  Hence, after substituting into our expression and simplifying using the identity
$$\sum_{i=1}^{n-1}\sum_{j=i+1}^{n}v_i v_j~=~\frac{1}{2} \left( \sum_{i=1}^{n} \right)^2-\frac{1}{2} \sum_{i=1}^{n} v_i^2,$$
we arrive at the following upper bound on robustness in terms of the connected components:
\begin{align}
R_{\mathrm{max}}(C_1, \ldots C_n) =
\frac{\frac{1}{2} ((\sum_i n_i)^2 - \sum_i n_i^2) + \sum_i \connectivity_i R(C_i)}
{\frac{1}{2} ((\sum_i v_i)^2 - \sum_i v_i^2) + \sum_i \connectivity_i}, \label{eq:exsccmaxrobustness}
\end{align}
Furthermore, this bound is attained.  Suppose that $G_{\mathrm{tot}}$ is the graph on $n$ nodes with a link from node $i$ to node $j$ whenever $i < j$.  Then, by the construction described in the previous section on network hierachy, we have a graph $G_{\mathrm{max}}$ such that the components of $G_{\mathrm{max}}$ are $C_1, \ldots C_n$ and $\hier (G_{\mathrm{max}}) = G_{\mathrm{tot}}$.  By our formula, $R(G_{\mathrm{max}}) = R_{\mathrm{max}}(C_1, \ldots C_n)$.
